---
title: "General Social Survey"
author: "Mykela Castillejo"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```


```{r}
barchartGC(~VOTE16+VOTE12, data=mydata4, type = "percent") 
chisqtestGC(~VOTE16+VOTE12, data=mydata4)
votetab=xtabs(~VOTE16+VOTE12, data=mydata4)
fisher.test(votetab)
```

# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. First explain what you make out of comparing how people voted in 2012 with 2016.


## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

By using variable analysis we will see the relationship between those that believe in the bigbang theory and those that believe in the afterlife.Then I will do a chai-square test. The barchart will show the differences between the two subjects. Then when the fisher test will provides the numerical results.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
voted1=c(92.04 , 35.96)
voted2=c(35.96 , 14.04)
tabl = rbind(voted1,voted2)
barchartGC(tabl)
```


#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
mytab<- xtabs(~VOTE16+VOTE12, data=mydata4)
rowPerc(mytab) 
colPerc(mytab)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

We reject the null hypothesis.

$H_0$ :There is no relationship between those that voted in 2016 and 2012  
$H_a$ :The ones that voted in 2016 and 2012 are related

```{r}
chisqtestGC(~VOTE16+VOTE12, data=mydata4)

```

For the fisherexact test we have 

$H_0$ : the odds of the ones that voted in 2016 and 2012 are one to one       
$H_a$ : the odds of the ones that voted in 2016 and 2012 do not contain number one

```{r}
fisher.test(mytab)
```

# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

By using variable analysis we will see the relationship between those that voted in 2016 and 2012.Then I will do a chai-square test. The barchart will show the differences between the two subjects. Then when the fisher test will provides the numerical results.


##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~BIGBANG+POSTLIFE, data=mydata4, type = "percent") 
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata4)
votetab=xtabs(~BIGBANG+POSTLIFE, data=mydata4)
fisher.test(votetab)
```
#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
mytab<- xtabs(~BIGBANG+POSTLIFE, data=mydata4)
rowPerc(mytab) 
colPerc(mytab)
```

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

We fail to reject the null hypothesis.

$H_0$ :There is no relationship between those that believe in the bigbang and those that believe that in the afterlife  
$H_a$ :Belief in bigbang and the afterlife are related

```{r}
chisqtestGC(~BIGBANG+POSTLIFE, data=mydata4)

```

For the fisherexact test we have 

$H_0$ : the odds of believing in the bigbang and the afterlife are one to one     
$H_a$ : the odds of believing in the bigbang and the afterlife do not contain number one

```{r}
fisher.test(mytab)
```

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This time compare the opinion on death penalty (CAPPUN) with opinion on courts (COURTS).

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

By using variable analysis we will see the relationship between  the opinion on death penalty with opinion on courts. Then I will use the data in the chai-square test.The barchart will show the differences between the two subjects. Then when the fisher test will provides the numerical results.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN+COURTS, data=mydata4, type = "percent") 
chisqtestGC(~CAPPUN+COURTS, data=mydata4)
votetab=xtabs(~CAPPUN+COURTS, data=mydata4)
fisher.test(votetab)
```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
mytab<- xtabs(~CAPPUN+COURTS, data=mydata4)
rowPerc(mytab) 
colPerc(mytab)
```


### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

We fail to reject the null hypothesis.

$H_0$ :There is no relationship between those that opinion on the death penalty and courts    
$H_a$ :The opinion on the death penalty and courts are related

```{r}
chisqtestGC(~CAPPUN+COURTS, data=mydata4)

```

For the fisherexact test we have 

$H_0$ : The opinion on the death penalty and courts are one to one       
$H_a$ : The opinion on the death penalty and courts do not contain number one

```{r}
fisher.test(mytab)
```


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

After reviewing the data of the three questions the first question has no relationship it stays the same. For the second question I have found that there is relationship because the p-value is between 1 and zero. The third question has a relationship the p-value is also below zero. 